## Trend analysis

For each location and litter type, the following statistics will be estimated for the period `r str_c(DATE_FROM, " to ", DATE_TO)`:

- from: the first date of the time-series;
- to: the final date of the time-series;
- [Theil-Sen slope](https://en.wikipedia.org/wiki/Theil%E2%80%93Sen_estimator) (slope): a robust non-parametric estimator of slope (counts / year);
- [p-value](https://en.wikipedia.org/wiki/P-value): the p-value associated with the one-tailed [Mann-Kendall test](https://en.wikipedia.org/wiki/Kendall_rank_correlation_coefficient) to test the null hypothesis of 
    - no monotonically _increasing_ trend in case the Theil-Sen slope is greater than zero;
    - no monotonically _decreasing_ trend in case the Theil-Sen slope is smaller than zero;
- the number of surveys (N);

The Mann-Kendall test is a non-parametric test and as such does not make distributional assumptions on the data. 

<br/>
<br/>

```{r}
# create table with summary statistics
d_stats %>%
    filter(type_code %in% TYPE_CODES) %>%
    group_by(type_name) %>%
    mutate(mean_ranking = mean(ranking)) %>%
    ungroup %>%
    arrange(desc(is_group_code(type_code)), mean_ranking, region_name, country_code, location_name) %>%
    mutate(
        p_value = formatC(p_value, format = "f", digits = 3),
        slope = b1 * 365.25) %>%
    select(country_code, location_name, type_name, from, to, slope, `p-value` = p_value, N = n) %>%
    mutate_if(
        is_double,
        formatC, format = "fg", digits = 3) %>%
    kable(align = "lllllrrr")
```


<p><br/></p>


Time-series of the selected litter types are given in the plots below, including trend line and smoother.

The lines and dots have the following meaning:

- dots: observations;
- thin gray line segments: auxiliary line segments to guide the eye (only given for 5 or more points);
- blue line: [loess-smoother](https://en.wikipedia.org/wiki/Local_regression). It is only given for 18 or more points;
- red line: Theil-Sen trend line (its slope is given in the table above). It is only given for 5 or more points.

> Descriptive statisics and trend analysis results have been stored in file `r sQuote(basename(STATS_FILE_NAME))`.


```{r fig.width=6, fig.height=4, out.width=730}
d_ltr %>%
    select(region_name, country_code, location_name, date, type_name, abundance) %>%
    left_join(d_stats, by = c("region_name", "country_code", "location_name", "type_name")) %>%
    arrange(ranking) %>%
    split(group_indices(., region_name, country_code, location_name, type_name)) %>%
    map_df(function(x) {
        g <- ggplot(data = x, mapping = aes(x = date, y = abundance)) +
                geom_point() +
                scale_x_date(name = "", limits = c(DATE_FROM, DATE_TO)) +
                ggtitle(str_c(x$location_name[1], x$type_name[1], sep = "  "))
        if (nrow(x) >= 5L) {
            if (nrow(x) <= 24L) {
                g <- g +        
                    geom_path(size = 0.5, colour = "gray")
            }
            d <- tibble(date = range(x$date)) %>%
                mutate(abundance = x$b0[1] + x$b1[1] * as.numeric(date))
            g <- g + 
                geom_point() +
                geom_path(
                    data = d, 
                    mapping = aes(x = date, y = abundance),
                    colour = "red", size = 1.0)
            if (nrow(x) >= 18L) {
                g <- g +        
                    geom_smooth(method = "loess", se = FALSE, colour = "blue", size = 0.5)
            }
        }
        tibble(
            g = list(g), 
            ranking = x$ranking[1], 
            region_name = x$region_name[1],
            country_code = x$country_code[1],
            location_name = x$location_name[1], 
            type_name = x$type_name[1])
    }) %>%
    arrange(desc(is_group_code(type_name)), location_name, ranking, type_name) %>%
    pull("g") %>%
    walk(print)
```