---
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r}
logger$info("Entering regional trend analysis section")
```

<br/>
<br/>

## Regional trend analysis

For each region_code and the type names and group codes specified in the settings file, the following statistics have been estimated for the period `r str_c(DATE_FROM, " to ", DATE_TO)`:

- the number of surveys (N);
- Theil-Sen slope: the median of all [Theil-Sen slopes](https://en.wikipedia.org/wiki/Theil%E2%80%93Sen_estimator) within a region;
- [p-value](https://en.wikipedia.org/wiki/P-value): the p-value associated with the one-tailed Regional Kendall test ([Van Belle & Hughes, 1984](http://dx.doi.org/10.1029/WR020i001p00127); [Gilbert, 1987](https://www.osti.gov/biblio/7037501-statistical-methods-environmental-pollution-monitoring)) to test the null hypothesis of 
    - no monotonically _increasing_ trend in case the regional Theil-Sen slope is greater than zero;
    - no monotonically _decreasing_ trend in case the regional Theil-Sen slope is smaller than zero;

A p-value less than an _a priori_ specified [significance level](https://en.wikipedia.org/wiki/Statistical_significance) (_e.g._, often &alpha; = 0.05), indicates a significant trend. If the p-value is greater than this significance level, we can't say that there is no trend. We can only conclude that our data do not show evidence for a significant trend (due to lack of data, noise, etc.).

The Regional Kendall test is a non-parametric test and as such does not make distributional assumptions on the data.

Note that the trend statistics can only be computed if all `location_code`s of a `region_code` have at least three records (surveys). If that is not the case, the table below contains `NA`.

<br/>
<br/>

```{r}
# check region_codes
missing_region_codes <- REGION_CODE %>%
  setdiff(unique(d_stats_reg$region_code))
if (length(missing_region_codes) > 0) {
    logger$warn("The following specified region code(s) are not found and will be skipped: ",
        missing_region_codes %>% sQuote %>% enumerate)
}
REGION_CODE <- d_stats_reg %>%
    chuck("region_code") %>% 
    unique %>%
    intersect(REGION_CODE)

# check type names
missing_type_names <- TYPE_NAME %>%
  setdiff(unique(d_stats_reg$type_name))
if (length(missing_type_names) > 0) {
    logger$warn("The following specified type name(s) are not found and will be skipped: ",
        missing_type_names %>% sQuote %>% enumerate)
}
TYPE_NAME <- d_stats_reg %>%
    chuck("type_name") %>% 
    unique %>%
    intersect(TYPE_NAME)

# check group codes
missing_group_codes <- GROUP_CODE %>%
  setdiff(unique(d_stats_reg$type_name))
if (length(missing_group_codes) > 0) {
    logger$warn("The following specified group code(s) are not found and will be skipped: ",
        missing_group_codes %>% sQuote %>% enumerate)
}
GROUP_CODE <- d_stats_reg %>%
    chuck("type_name") %>% 
    unique %>%
    intersect(GROUP_CODE)
```

```{r}
logger$info("Creating table with trend statistics")
d_stats_reg %>%
    filter(type_name %in% unique(c("TC", TYPE_NAME, GROUP_CODE))) %>%
    select(region_code, type_name, N = n, slope, `p-value` = p_value) %>%
    arrange(region_code, desc(abs(slope))) %>%
    rename(`type name / group code` = type_name) %>%
    mutate(
        slope = slope %>%
            formatC(format = "fg", digits = 4),
        `p-value` = `p-value` %>%
            formatC(format = "f",  digits = 4)) %>%
    kable(align = "llrrr")
logger$info("Table with trend statistics created")
```

<br/>
<br/>
```{r eval=length(REGION_CODE) == 0L}
  logger$info("No valid region_code(s) found. Therefore, time-series plot are skipped")
```

```{r, child=system.file("app", "litter-trend-plots-regional.Rmd", package = "litteR"), eval=length(REGION_CODE) != 0L}
```