---
title:  | 
        | <font color="blue"> LitteR - Analysis of Litter Data</font>
        | <font size="1"><i> &copy; 2018-2019 D.J.J. Walvoort & W.M.G.M Van Loon. [MIT-license](https://opensource.org/licenses/MIT). `r sprintf("litteR version: %s (%s)", packageVersion("litteR"), packageDate("litteR"))`</i></font>
author: |
        | Dennis Walvoort (<font size="1">WENR-WUR, The Netherlands</font>) Willem van Loon (<font size="1">RWS, The Netherlands</font>)
date:   | 
        | `r format(Sys.time())`
        | 
        |
pagetitle: "litteR-report"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: false
editor_options: 
  chunk_output_type: console
params:
  analyst_name:
    label: "Your name:"
    value: "not specified"
    placeholder: "Enter your name"
    input: text
  file_stats:
    label: "Name report:"
    value: ""
    placeholder: "Enter name of stats output file"
    input: file
  file_input: 
    label: "Input file:"
    value: ""
    placeholder: "Enter name of litter file"
    input: file
  file_groups: 
    label: "Group file:"
    value: ""
    placeholder: "Enter name of file with group definitions"
    input: file
  module_stats:
    label: "Descriptive statistics"
    value: TRUE
    input: checkbox
  module_trend:
    label: "Trend analysis"
    value: TRUE
    input: checkbox
  module_baseline:
    label: "Baseline analysis"
    value: TRUE
    input: checkbox
  module_power:
    label: "Power analysis"
    value: FALSE
    input: checkbox
  percentage_total_abundance: 
    label: "Percentage of total abundance:"
    value: 80
    min: 50
    max: 100 
    step: 1
    width: 200
    input: numeric
  min_date:
    label: "First date to analyse"
    value: "2011-01-01"
    startview: year
    min: "1990-01-01"
    max: "2050-12-31"
    format: "yyyy-mm-dd"
    width: 200
    input: date
  max_date:
    label: "Final date to analyse"
    value: "2017-12-31"
    startview: year
    min: "1990-01-01"
    max: "2050-12-31"
    format: "yyyy-mm-dd"
    width: 200
    input: date
  number_of_simulations: 
    label: "Number of simulations:"
    value: 100
    min: 100
    max: 10000
    step: 100
    width: 200
    input: numeric
  alpha: 
    label: "significance level:"
    value: 0.05
    min: 0.001
    max: 0.1
    step: 0.001
    width: 200
    input: numeric
  resolution_effect_size: 
    label: "Resolution effect size:"
    value: 10
    min: 1
    max: 50
    step: 1
    width: 200
    input: numeric
  min_surveys:
    label: "Minimum number of surveys"
    value: 24
    min: 16
    max: 100
    step: 1
    width: 200
    input: numeric
  litter_types:
    label: Litter types
    value: [[TA], [49]]
    choices: ['[TA]', '[15]','[19]','[45]']
    input: select
    multiple: TRUE
  show_source_code:
    label: Show source code?
    value: FALSE
    input: checkbox
  image_quality:
    label: "image quality"
    value: "high"
    input: text
---



```{r include=FALSE, echo=params$show_source_code}
knitr::opts_knit$set(
  progress = FALSE, 
  verbose = FALSE
)

# make sure all table columns are visible
options(dplyr.width = Inf, warning.length = 8000)
```



```{r, include=FALSE, message=FALSE, warning=FALSE, echo=params$show_source_code}

# Coding convention
# To make scripts readable and consistent, we adhere to the packages and 
# programming style of the tidyverse (www.tidyverse.org)

# attach packages to R's search path
suppressPackageStartupMessages({
  library(rlang)     # tidyverse versions of base/core functions
  library(dplyr)     # grammar of data manipulation
  library(purrr)     # functional programming tools
  library(tidyr)     # tidy data
  library(ggplot2)   # grammar of graphics
  library(fs)        # file system
  library(knitr)     # dynamic report generation
  library(litteR)    # litter analysis
  library(readr)     # writing CSV-files
  library(stringr)   # common string operations
})
```



```{r, echo=params$show_source_code}
# image quality
IMAGE_QUALITY <- params %>%
  chuck("image_quality")
permissible_image_quality <- c(low = 72, high = 300)
if (!(IMAGE_QUALITY %in% names(permissible_image_quality))) {
    warning(
        "image quality should be set to either %s or %s" %>%
            sprintf(
                names(permissible_image_quality)[1], 
                names(permissible_image_quality)[2]),
        call. = FALSE)
    message("Assuming publication quality")
    IMAGE_QUALITY <- "publication"
}

# visibility of source code
SHOW_SOURCE_CODE <- params %>%
  pluck("show_source_code", .default = FALSE)
```


```{r include=FALSE, purl=FALSE}
# global knitr-settings (defaults)
opts_chunk$set(
    echo = SHOW_SOURCE_CODE,
    comment = NULL,
    dpi = c(low = 72, high = 300)[IMAGE_QUALITY]
)
```




```{r, echo=FALSE}

# analyst's name
ANALYST_NAME <- params %>%
  pluck("analyst_name", .default = "unknown")

# files
INPUT_FILE_NAME <- params %>%
  chuck("file_input")
GROUPS_FILE_NAME <- params %>%
  chuck("file_groups")
STATS_FILE_NAME <- params %>%
  chuck("file_stats")
META_FILE_NAME <- STATS_FILE_NAME %>%
  str_replace("stats", "stats-meta")

# tasks to perform
MODULE_STATS <- params %>% 
  pluck("module_stats", .default = FALSE)
MODULE_TREND <- params %>% 
  pluck("module_trend", .default = FALSE)
MODULE_BASELINE <- params %>% 
  pluck("module_baseline", .default = FALSE)
MODULE_POWER <- params %>% 
  pluck("module_power", .default = FALSE)

# period to analyse (DATE_FROM <= DATE_TO)
DATE_FROM <- params$min_date %>% as.Date
DATE_TO <- params$max_date %>% as.Date

# percentage of total abundance
PERCENTAGE_TOTAL_ABUNDANCE <- params %>%
  pluck("percentage_total_abundance", .default = 80)

# litter types to analyse
TYPE_CODES <- params %>% 
  chuck("litter_types") %>%
  map_chr(str_to_upper) %>%
  unique

# significance level
ALPHA <- params %>%
  chuck("alpha")

# number of Monte Carlo simulations
N_SIM <- params %>%
  chuck("number_of_simulations")

# resolution of effect size power analysis
RESOLUTION_EFFECT_SIZE <- params %>%
  chuck("resolution_effect_size")

# minimum number of surveys to sample from
MIN_SURVEYS <- params %>%
  chuck("min_surveys")
```


```{r}
# read and validate group file
d_grp <- read_litter_groups(GROUPS_FILE_NAME)  %>%
  mutate(group_name = str_c("[", group_name, "]"))
```


### Introduction

This tool performs the following types of data analysis on litter data, or any other type of litter data: data quality control, exploratory data analysis, trend analysis, baseline analysis and power analysis.

This report can best be viewed with the latest versions of web browsers like Google Chrome, Mozilla Firefox, Chromium, or Safari. Its contents does not render well in some versions of Microsoft's Internet Explorer.



### Settings

- analyst: `r ANALYST_NAME`
- files:
    - project path: `r sQuote(path_dir(INPUT_FILE_NAME))`
    - input file: `r sQuote(path_file(INPUT_FILE_NAME))`
    - input file signature: `r sQuote(get_signature(path(INPUT_FILE_NAME)))`
    - group file: `r sQuote(path_file(GROUPS_FILE_NAME))`
- period: from `r DATE_FROM` to `r DATE_TO`
- modules:
    - descriptive statistics: `r str_to_lower(MODULE_STATS)`
    - trend analysis: `r str_to_lower(MODULE_TREND)`
    - baseline analysis: `r str_to_lower(MODULE_BASELINE)`
    - power analysis: `r str_to_lower(MODULE_POWER)`
- litter type(s): `r TYPE_CODES`
- exploratory data analysis
    - percentage of total abundance to analyse: `r PERCENTAGE_TOTAL_ABUNDANCE`%
- power analysis
    - significance level (&alpha;): `r ALPHA`
    - number of Monte Carlo simulations: `r N_SIM`
    - minimum surveys to sample from: `r MIN_SURVEYS`
    - resolution of the effect size: `r RESOLUTION_EFFECT_SIZE`%
- report
    - image quality: `r sQuote(IMAGE_QUALITY)`






```{r various checks, echo=FALSE, results='hide'}

(INPUT_FILE_NAME == "") && stop(
  "Litter file not specified",
  call. = FALSE
)

!file_exists(INPUT_FILE_NAME) && stop(
    sprintf("Litter file %s not found", sQuote(INPUT_FILE_NAME)),
    call. = FALSE
)

(DATE_FROM > DATE_TO) && stop(
    "first date '%s' is later than final date '%s'" %>%
        sprintf(DATE_FROM, DATE_TO),
    call. = FALSE
)

```

```{r, eval=MODULE_POWER, results='hide'}
((ALPHA < 0.001) | (ALPHA > 0.1)) && stop(
    "ALPHA should be in the range from 0.001 to 0.1",
    call. = FALSE
)

(!is.integer(N_SIM) || (N_SIM < 0)) && stop(
    "N_SIM should be a positive integer",
    call. = FALSE
)

((RESOLUTION_EFFECT_SIZE < 1) | (RESOLUTION_EFFECT_SIZE > 50)) && stop(
    "Step effect size for power analysis should be in the range from 1 to 50%",
    call. = FALSE
)

(!is.integer(MIN_SURVEYS) || (MIN_SURVEYS < 15)) && stop(
    "MIN_SURVEYS should be a positive integer greater than or equal to 15",
    call. = FALSE
)

```


<br/>
<br/>

### Data Quality Control

In this section, litter data will be read and validated. Warnings will be printed if they occur.


```{r}
# read and validate input file
d_all <- INPUT_FILE_NAME %>%
  read_litter %>%
  select(-country_name) %>%
  filter(date %>% between(DATE_FROM, DATE_TO))

(nrow(d_all) == 0) && stop(
    "No data available between '%s' and '%s'.\nPlease check '%s' and '%s'." %>%
        sprintf(DATE_FROM, DATE_TO, path_file(INPUT_FILE_NAME), "settings.yaml"),
    call. = FALSE
)

```


```{r}
# add percentage and top-X indicator
d_top <- d_all %>%
    group_by(region_name, country_code, location_code, location_name, type_name) %>%
    summarise(total_abundance_litter = sum(abundance)) %>%
    mutate(total_abundance_location_name = sum(total_abundance_litter)) %>%
    mutate(perc_abundance = 100 * total_abundance_litter / total_abundance_location_name) %>%
    arrange(location_name, desc(perc_abundance)) %>%
    mutate(
        cum_perc = cumsum(perc_abundance), 
        topx = cum_perc < (PERCENTAGE_TOTAL_ABUNDANCE + 0.001)) %>%
    ungroup %>%
    select(region_name, country_code, location_code, location_name, type_name, perc_abundance, topx)

# add ranking
d_top <- d_top %>%
    group_by(region_name, country_code, location_code, location_name) %>%
    mutate(ranking = rank(-perc_abundance, ties.method = "first")) %>%
    ungroup %>%
    arrange(region_name, country_code, location_code, location_name, ranking)
```



```{r results='hide'}

# get user-specified codes in settings file
specified_codes <- TYPE_CODES %>% 
    get_type_code %>%
    discard(are_na)

# get user-specified codes in settings file
specified_groups <- TYPE_CODES %>% 
    get_group_code %>%
    discard(are_na)

# get available codes in data file
available_codes_data_file <- d_all %>%
    pull("type_name") %>%
    get_type_code %>%
    unique

# get available groups in group file
available_codes_group_file <- d_grp %>%
    pull("type_name") %>%
    get_type_code %>%
    unique

# get available groups in group file
available_groups_group_file <- d_grp %>%
    pull("group_name") %>%
    unique

# check if all type codes as specified in settings.yaml also occur in the data file
all(specified_codes %in% available_codes_data_file) || stop(sprintf(str_c(
    "The following type code(s) are specified as %s in file %s ",
    "but missing in input file %s:\n%s"),
    "litter_types" %>%
      sQuote %>%
      enumerate,
    "settings.yaml" %>%
      sQuote,
    INPUT_FILE_NAME %>%
      path_file %>%
      sQuote,
    specified_codes %>%
      setdiff(available_codes_data_file) %>%
      enumerate,
    call. = FALSE
))

# check if all group codes as specified in settings.yaml also occur in the group file
all(specified_groups %in% available_groups_group_file) || stop(sprintf(str_c(
    "The following groups are specified in file %s ",
    "but missing in group file %s:\n%s"),
    "settings.yaml" %>%
      sQuote,
    GROUPS_FILE_NAME %>%
      path_file %>%
      sQuote,
    specified_groups %>%
      setdiff(available_groups_group_file) %>%
      enumerate,
    call. = FALSE
))

# check if all type codes in data file are also in group file
all(available_codes_data_file %in% available_codes_group_file) || stop(sprintf(str_c(
    "The following types in data file %s are missing in group file %s:\n%s\n",
    "Please add these types to the group file and indicate the groups they belong to."),
    INPUT_FILE_NAME %>%
      path_file %>%
      sQuote,
    GROUPS_FILE_NAME %>%
      path_file %>%
      sQuote,
    available_codes_data_file %>%
      setdiff(available_codes_group_file) %>%
      enumerate,
    call. = FALSE
))

# check if all type codes in data file are also in group file
if (!all(available_codes_group_file %in% available_codes_data_file)) {
  warning(sprintf(
    "The following types in group file %s are missing in data file %s:\n%s",
    GROUPS_FILE_NAME %>%
      path_file %>%
      sQuote,
    INPUT_FILE_NAME %>%
      path_file %>%
      sQuote,
    available_codes_group_file %>%
      setdiff(available_codes_data_file) %>%
      enumerate,
    call. = FALSE
  )) 
}

```


```{r}
# add litter groups to litter types
d_all <- d_grp %>%
  pull("group_name") %>%
  unique %>%
  map_df(function(x) {
    d_grp %>%
      filter(group_name == x) %>%
      left_join(d_all, by = "type_name") %>%
      group_by(region_name, country_code, location_code, location_name, date) %>%
      summarise(abundance = sum(abundance)) %>%
      ungroup %>%
      filter(!are_na(abundance)) %>%
      mutate(type_name = x)
  }) %>%
  bind_rows(d_all)

# order by abundance
d_all <- d_all %>% 
    left_join(d_top, by = c("region_name", "country_code", "location_code", "location_name", "type_name")) %>%
    mutate(
      topx = topx %>% coalesce(FALSE),
      ranking = ranking %>% coalesce(0L)) %>%
    arrange(region_name, country_code, location_code, location_name, ranking, date)

# add codes
d_all <- d_all %>%
    mutate(
      type_code = get_type_code(type_name),
      type_code = if_else(are_na(type_code), get_group_code(type_name), type_code))
    
```


```{r}
# Select type(s) of interest
d_ltr <- d_all %>%
    filter(type_code %in% TYPE_CODES)
```

```{r}
d_stats <- d_all %>%
    group_by(region_name, country_code, location_code, location_name, type_name, type_code) %>%
    summarise(
        from = format(min(date)),
        to = format(max(date)),
        mean = mean(abundance),
        median = median(abundance),
        cv = cv(abundance),
        rmad = rmad(abundance),
        n = n(),
        ts = list(theil_sen(date, abundance)),
        b0 = ts %>%
          first %>%
          intercept,
        b1 = ts %>%
          first %>%
          slope,
        p_value = if_else(
            are_na(b1),
            NA_real_,
            if_else(
                b1 >  0,
                p_value(mann_kendall(abundance, "increasing")),
                p_value(mann_kendall(abundance, "decreasing"))
            )
        ),
        min = min(abundance),
        p01 = quantile(abundance, 0.01),
        p05 = quantile(abundance, 0.05),
        p10 = quantile(abundance, 0.10),
        p25 = quantile(abundance, 0.25),
        p50 = quantile(abundance, 0.50),
        p75 = quantile(abundance, 0.75),
        p90 = quantile(abundance, 0.90),
        p95 = quantile(abundance, 0.95),
        p99 = quantile(abundance, 0.99),
        max = max(abundance)
        ) %>%
    ungroup %>%
    select(-ts) %>%
    left_join(d_top, by = c("region_name", "country_code", "location_code", "location_name", "type_name")) 

# add relative abundance to groups
d_stats <- d_stats %>% 
    split(., group_indices(., region_name, country_code, location_code, location_name)) %>%
    map_df(function(x) {
      ta <- x$mean[x$type_code == "[TA]"]
      sel <- are_na(x$perc_abundance)
      x$perc_abundance[sel] <- 100 * x$mean[sel] / ta
      x
    })


# store stats in a CSV-file
d_stats %>%
    rename(intercept = b0, slope = b1, p_value_slope = p_value) %>%
    mutate(slope = slope * 365.25) %>%
    select(-perc_abundance, -topx, -ranking) %>%
    mutate_if(is_double, function(x){if_else(is.nan(x), NA_real_, x)}) %>%
    mutate_if(is_double, signif, digits = 3) %>%
    write_csv(STATS_FILE_NAME, na = "")

# store corresponding metadata
path_package("litteR", "extdata", "litter-stats-meta.csv") %>%
  read_csv(
    col_types = cols(
      column_name = col_character(),
      description = col_character(),
      unit = col_character()
    )
  ) %>%
  write_csv(META_FILE_NAME, na = "")
```



<br/>
<br/>
```{r, child=system.file("app", "litter-stats.Rmd", package = "litteR"), eval=MODULE_STATS}
```

<br/>
<br/>
```{r, child=system.file("app", "litter-trend.Rmd", package = "litteR"), eval=MODULE_TREND}
```

<br/>
<br/>
```{r, child=system.file("app", "litter-baseline.Rmd", package = "litteR"), eval=MODULE_BASELINE}
```

<br/>
<br/>
```{r, child=system.file("app", "litter-power.Rmd", package = "litteR"), eval=MODULE_POWER}
```


## Session Information

```{r}
sessionInfo()
```

