---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
logger$info("Entering descriptive statistics section")
```

<br/>
<br/>

## Descriptive statistics

```{r, echo=FALSE}
MIN_YEARS <- 5L
MIN_SURVEYS <- 10L
```

The number of years and the number of surveys for each `location_code` should not be too small, otherwise the calculations in this report will be less reliable. In addition, the surveys should ideally also be evenly spread in time.

Note that **litteR** does not enforce a minimum number of years or surveys. That is the responsibility of the user. As a guideline, we advise a minimum of `r MIN_YEARS` years and `r MIN_SURVEYS` surveys, evenly distributed in time.

The table below gives the number of surveys and the number of years for each `location_code`.

<br/>
<br/>

```{r}

if (REGIONAL_ANALYSIS) {
    d <- d_ltr %>%
        filter(type_name == "TC") %>%
        transmute(
            region_code,
            location_code,
            date,
            year = date %>%
                format("%Y"))  %>%
        group_by(region_code, location_code) %>%
        summarise(
            `number of years` = length(unique(year)),
            `number of surveys` = n(),
            .groups = "drop") 
    d %>%
        kable(align = "llrr")
} else {
    d <- d_ltr %>%
        filter(type_name == "TC") %>%
        transmute(
            location_code,
            date,
            year = date %>%
                format("%Y"))  %>%
        group_by(location_code) %>%
        summarise(
            `number of years` = length(unique(year)),
            `number of surveys` = n(),
            .groups = "drop") 
    d %>%
        kable(align = "lrr")
}
```

<br/>

```{r}
if (any(d %>% chuck("number of years") < MIN_YEARS)) {
    location_code <- d %>%
        filter(`number of years` < MIN_YEARS) %>%
        chuck("location_code")
    logger$warn(
        str_glue("The number of years may be insufficient (<{MIN_YEARS} years) for: "),
        location_code %>% sQuote %>% enumerate)
}    
if (any(d %>% chuck("number of surveys") < MIN_SURVEYS)) {
    location_code <- d %>%
        filter(`number of surveys` < MIN_SURVEYS) %>%
        chuck("location_code")
    logger$warn(
        str_glue("The number of surveys may be insufficient (<{MIN_SURVEYS} surveys) for: "),
        location_code %>% sQuote %>% enumerate)
}    
if (any(d %>% chuck("number of surveys") < 3L)) {
    location_code <- d %>%
        filter(`number of surveys` < 3L) %>%
        chuck("location_code")
    logger$warn(
        "The number of surveys is insufficient (<3 surveys) to estimate the regional trend for: ",
        location_code %>% sQuote %>% enumerate)
}    

```


<br/>
<br/>


For each location code and group/type name, the following statistics have been estimated for the period `r str_c(DATE_FROM, DATE_TO, sep = " to ")`:

-   mean count (`mean`):, *i.e.*, [the arithmetic mean](https://en.wikipedia.org/wiki/Arithmetic_mean) of the counts for each litter type;

-   median count (`median`), *i.e.*, the [median](https://en.wikipedia.org/wiki/Median) of the counts for each litter type;

-   relative count (`%TC`): the contribution of each litter type to the total count of litter types (%);

-   [coefficient of variation](https://en.wikipedia.org/wiki/Coefficient_of_variation) (`cv`): the ratio of the standard deviation to the mean of the counts for each litter type (%);

-   ratio of [mad](https://en.wikipedia.org/wiki/Median_absolute_deviation) and [median](https://en.wikipedia.org/wiki/Median) (`rmad`, %);

-   number of surveys (`n`);

-   [Theil-Sen slope](https://en.wikipedia.org/wiki/Theil%E2%80%93Sen_estimator) (slope): a robust non-parametric estimator of slope (litter counts / year);

-   [p-value](https://en.wikipedia.org/wiki/P-value): the p-value associated with the one-tailed [Mann-Kendall test](https://en.wikipedia.org/wiki/Kendall_rank_correlation_coefficient) to test the null hypothesis of

    -   no monotonically *increasing* trend in case the Theil-Sen slope is greater than zero;
    -   no monotonically *decreasing* trend in case the Theil-Sen slope is smaller than zero;

These statistics will be estimated for types with the greatest counts making up `r PERCENTAGE_TOTAL_COUNT`% of the total count for each location and for all groups specified in `r sQuote(path_file(FILE_TYPES))`.

> These statistics have been stored in file `r sQuote(path_file(FILE_STATS))`.

```{r}
logger$info("Creating table with litter statistics")
d_stats <- d_ltr %>%
    group_by(location_code, type_name, topx) %>%
    summarise(
        from = format(min(date)),
        to = format(max(date)),
        rel_count = mean(rel_count),
        mean = mean(count),
        median = median(count),
        cv = cv(count),
        rmad = rmad(count),
        n = n(),
        ts = list(theil_sen(date, count)),
        b0 = ts %>%
          first %>%
          intercept,
        b1 = ts %>%
          first %>%
          slope,
        slope = b1 * 365.25,
        p_value = if_else(
            are_na(b1),
            NA_real_,
            if_else(
                b1 > 0,
                p_value(mann_kendall(count, type = "increasing")),
                p_value(mann_kendall(count, type = "decreasing"))
            )
        ),
        .groups = "drop") %>%
    select(-ts)

# store stats in a CSV-file
d <- d_stats %>%
    filter(topx | are_na(topx)) %>% # only top x% or groups
    select(location_code, from, to, type_name,
           rel_count, mean, median, cv, rmad, n, slope, p_value) %>%
    arrange(location_code, desc(rel_count)) %>%
    rename(`%TC` = rel_count)

d %>%
    rename(`type/group_name` = type_name) %>%
    mutate(across(where(is_double) & !p_value, ~ formatC(.x, format = "fg", digits = 4))) %>%
    mutate(p_value = formatC(p_value, format = "f", digits = 4)) %>%
    mutate(across(where(is_character), ~ str_remove(.x, "^ +"))) %>%
    write_csv(FILE_STATS)

logger$info("Table with litter statistics created")
```

The statistics for the litter groups are given in the table below.

<br/>
<br/>

```{r}
d %>%
    filter(type_name == str_to_upper(type_name)) %>% # group codes
    rename(group_code = type_name) %>%
    mutate(across(where(is_double) & !p_value, ~ formatC(.x, format = "fg", digits = 4))) %>%
    mutate(p_value = formatC(p_value, format = "f", digits = 4)) %>%
    kable(align = "llllrrrrrrrr")
```

<br/>
<br/>

The figures below show for each location code the median count for each group.

```{r fig.width=5, fig.height=5, out.width=449}
d_stats %>%
    select(location_code, type_name, median) %>%
    filter(type_name == str_to_upper(type_name)) %>% # group codes
    arrange(location_code) %>%
    split(.$location_code) %>%
    walk(function(x) {
        x$type_name <- x %>%
            chuck("type_name") %>%
            factor(
                levels = d_type %>%
                    chuck("group_code") %>%
                    levels %>%
                    rev,
                ordered = TRUE)
      g <- ggplot(data = x) +
        geom_col(mapping = aes(x = type_name, y = median), fill = "blue", alpha = 0.5) +
        scale_x_discrete(name = "") +
        scale_y_continuous(name = "median count") +
        coord_flip() +
        ggtitle(x$location_code[1])
      print(g)
    })
```