## Descriptive statistics

For each location and litter type, the following statistics will be estimated for the period `r str_c(DATE_FROM, DATE_TO, sep = " to ")`:

-  mean abundance (mean):, _i.e._, [the arithmetic mean](https://en.wikipedia.org/wiki/Arithmetic_mean) of the counts for each litter type;
- median abundance (median), _i.e._, the [median](https://en.wikipedia.org/wiki/Median) of the counts for each litter type;
- relative abundance (rel.abund.): the contribution of each litter type to the total abundance of litter types (%);
- [coefficient of variation](https://en.wikipedia.org/wiki/Coefficient_of_variation) (CV): the ratio of the standard deviation to the mean of the counts for each litter type (%);
- ratio of the [MAD](https://en.wikipedia.org/wiki/Median_absolute_deviation) and the [median](https://en.wikipedia.org/wiki/Median) (RMAD, %);
- number of surveys (N);

These statistics will be estimated for types with the greatest abundances making up `r PERCENTAGE_TOTAL_ABUNDANCE`% of the total abundance for each location.

> These statisics have been stored in file `r sQuote(basename(STATS_FILE_NAME))`.


```{r}

# aggregation to various spatial scales

d_assessment_location <- d_stats %>%
    filter(topx | type_name == type_code) %>%
    mutate(RMAD = rmad * 100, CV = cv * 100) %>%
    select(country_code, region_name, location_name, type_name, from, to,
           mean, median, `rel.abund.` = perc_abundance, CV, RMAD, N = n)

d <- d_assessment_location %>%
    select(-from, -to, -location_name)

d_assessment_country_region <- d %>%
    group_by(country_code, region_name, type_name) %>%
    summarise_all(mean) %>%
    ungroup

show_country_region_level <- (d %>%
    group_by(country_code, region_name) %>%
    summarise(n = n()) %>%
    nrow) > 1

d_assessment_country <- d %>%
    select(-region_name) %>%
    group_by(country_code, type_name) %>%
    summarise_all(mean) %>%
    ungroup 

d_assessment_regional <- d %>%
    select(-country_code) %>%
    group_by(region_name, type_name) %>%
    summarise_all(mean) %>%
    ungroup
```


Furthermore, these statistics are averaged for `r if_else(show_country_region_level, "four", "three")` spatial scales. These results are given below.


### Location level

#### Groups

```{r}
d_assessment_location %>%
    select(-country_code, -region_name) %>%
    filter(is_group_code(type_name)) %>%
    arrange(desc(`rel.abund.`)) %>%
    mutate_if(is_double, formatC, format = "fg", digits = 3) %>%
    kable(align = "llllrrrrrr")
```

#### Types

```{r}
d_assessment_location %>%
    select(-country_code, -region_name) %>%
    filter(!is_group_code(type_name)) %>%
    arrange(desc(`rel.abund.`)) %>%
    mutate_if(is_double, formatC, format = "fg", digits = 3) %>%
    kable(align = "llllrrrrrr")
```


```{r eval=show_country_region_level, results='asis'}
cat("### Country-regional level\n")
cat("#### Groups\n")
```

```{r eval=show_country_region_level}
d_assessment_country_region %>%
    filter(is_group_code(type_name)) %>%
    arrange(desc(is_group_code(type_name)), desc(`rel.abund.`)) %>%
    mutate_if(is_double, formatC, format = "fg", digits = 3) %>%
    kable(align = "lllrrrrrr")
```

```{r eval=show_country_region_level, results='asis'}
cat("#### Types\n")
```

```{r eval=show_country_region_level}
d_assessment_country_region %>%
    filter(!is_group_code(type_name)) %>%
    arrange(desc(is_group_code(type_name)), desc(`rel.abund.`)) %>%
    mutate_if(is_double, formatC, format = "fg", digits = 3) %>%
    kable(align = "lllrrrrrr")
```


### Country level

#### Groups

```{r}
d_assessment_country %>%
    filter(is_group_code(type_name)) %>%
    arrange(desc(is_group_code(type_name)), desc(`rel.abund.`)) %>%
    mutate_if(is_double, formatC, format = "fg", digits = 3) %>%
    kable(align = "llrrrrrr")
```

#### Types

```{r}
d_assessment_country %>%
    filter(!is_group_code(type_name)) %>%
    arrange(desc(is_group_code(type_name)), desc(`rel.abund.`)) %>%
    mutate_if(is_double, formatC, format = "fg", digits = 3) %>%
    kable(align = "llrrrrrr")
```


### Regional level

#### Groups

```{r message=FALSE}
d_assessment_regional %>%
    filter(is_group_code(type_name)) %>%
    arrange(desc(is_group_code(type_name)), desc(`rel.abund.`)) %>%
    mutate_if(is_double, formatC, format = "fg", digits = 3) %>%
    kable(align = "llrrrrrr")
```


#### Types

```{r message=FALSE}
d_assessment_regional %>%
    filter(!is_group_code(type_name)) %>%
    arrange(desc(is_group_code(type_name)), desc(`rel.abund.`)) %>%
    mutate_if(is_double, formatC, format = "fg", digits = 3) %>%
    kable(align = "llrrrrrr")
```